A Diminishing Payment Scheme of Employment Insurance Balances Financial Support and Reemployability for Insurance Recipients^[Code and data are available at; https://github.com/Mikmok14/UI-Spending-and-Income-Implications.git]

### 2022-02-27 outputs/paper/paper.Rmd
Add files via upload

---
title: A Diminishing Payment Scheme of Employment Insurance Balances Financial Support and Reemployability for Insurance Recipients^[Code and data are available at; https://github.com/Mikmok14/UI-Spending-and-Income-Implications.git]
author: "Marco Chau and Chyna Hui"
date: "18 February 2022"
output:
  bookdown::pdf_document2:
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
toc = FALSE:
abstract: This paper replicates the 2019 article *Consumer Spending during Unemployment Positive and Normative Implications*^[@mainrepref DOI:10.1257] which uses income, spending, and employment data to examine the normative implications for UI policy. This replication uses the original data provided by the authors and supplements it with external UI policy theories to evaluate the policy suggestions that the authors make. This replication finds that the UI pool is populated by myopic households that require an extension of financial support which are consistent with the original paper. This replication paper however finds that introducing a payment scheme of diminishing payments in combination with an extension of financial support duration leads to a more efficient market outcome.

@@

An initial observation of table \@ref(tab:spendchangetype) shows interesting prioritization of funds as income drops from UI exhaustion. Home improvement loses the most spending in terms of percentage. This is unsurprising as renovations projects can be put on hold until income becomes more stable and is not an immediate concern. Surprisingly, the drop in spending at discount stores, at -18.9% is more severe than department stores, at -17.7%. Intuitively, we would assume consumers would cut back on department store spending more than discount stores as income falls. Since income has decreased after UI has been exhausted, we expect consumers to switch to inferior goods which substitute normal goods reflected by a shift of spending from department stores to discount stores. This expectation is reflected from the transition of spending between Pre-UI and during UI periods where discount store spending went from \$57.70 to \$58.10 however the increase is quite marginal. The similar difference in food away from home at -15.9% and groceries at 15.8% suggests that households prioritize eating at home and eating out equally. The simultaenous reduction in both suggests that either a. households are in general eating the same quality of food but at a lesser amount or b. households are choosing cheaper alternatives to compensate for the loss of UI income.

@@

Juxtaposing the income and spending data, we both income and spending drop at similar times, however the degree of the reductions are quite different. In general, we observe that income drops much more than spending.

>>>>

---
title: A Diminishing Payment Scheme of Employment Insurance Balances Financial Support and Reemployability for Insurance Recipients^[Code and data are available at; https://github.com/Mikmok14/UI-Spending-and-Income-Implications.git]
author: "Marco Chau and Chyna Hui"
date: "18 February 2022"
output:
  bookdown::pdf_document2:
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
toc = FALSE:
abstract: This paper replicates the 2019 article *Consumer Spending during Unemployment Positive and Normative Implications*^[@mainrepref DOI:10.1257] which uses income, spending, and employment data to examine the normative implications for UI policy. This replication uses the original data provided by the authors and supplements it with external UI policy theories to evaluate the policy suggestions that the authors make. This replication finds that the UI pool is populated by myopic households that require an extension of financial support which is consistent with the original paper. This replication paper however finds that introducing a payment scheme of diminishing payments in combination with an extension of financial support duration leads to a more efficient market outcome.

@@

An initial observation of table \@ref(tab:spendchangetab) shows interesting prioritization of funds as income drops from UI exhaustion. Home improvement loses the most spending in terms of percentage. This is unsurprising as renovations projects can be put on hold until income becomes more stable and is not an immediate concern. Surprisingly, the drop in spending at discount stores, at -18.9% is more severe than department stores, at -17.7%. Intuitively, we would assume consumers would cut back on department store spending more than discount stores as income falls. Since income has decreased after UI has been exhausted, we expect consumers to switch to inferior goods which substitute normal goods reflected by a shift of spending from department stores to discount stores. This expectation is reflected from the transition of spending between Pre-UI and during UI periods where discount store spending went from \$57.70 to \$58.10 however the increase is quite marginal. The similar difference in food away from home at -15.9% and groceries at -15.8% suggests that households prioritize eating at home and eating out equally. The simultaneous reduction in both suggests that either a. households are in general eating the same quality of food but at a lesser amount or b. households are choosing cheaper alternatives to compensate for the loss of UI income.

@@

Juxtaposing the income and spending data, we see both income and spending drop at similar times, however the degree of the reductions are quite different. In general, we observe that income drops much more than spending.
